New insights into the Weddell Sea ecosystem applying a network approach

### 2022-11-23 Manuscript/Manuscript.tex
2nd version of MS

\begin{abstract}
The abstract goes here. It can also be on \emph{multiple lines}.
\end{abstract}

@@

The objective of this work was threefold: 1) estimate the strength for
each interaction in the Weddell Sea food web, 2) characterise species
considering weighted and unweighted properties, and 3) analyse the
species' role in the stability of the food web.

@@

The high Antarctic Weddell Sea shelf is situated between 74 and 78ºS
with a length of approximately 450 km (Figure 1). Water depth varies
from 200 to 500 m. Shallower areas are covered by continental ice, which
forms the coastline along the eastern and southern part of the Weddell
Sea. The shelf area contains a complex three-dimensional habitat with
large biomass, intermediate to high diversity in comparison to benthic
boreal communities and a spatially patchy distribution of organisms
\citep{Dayton1990, Teixido2002}.

@@

We obtained the dataset of the Weddell Sea food web from the GlobAL
daTabasE of traits and food Web Architecture (GATEWAy, version 1.0) of
the German Centre for Integrative Biodiversity Research (iDiv)
Halle-Jena-Leipzig \citep{Brose2018}. This open access database is a
list of predator-prey interactions that contains several highly-resolved
food webs, including biological data about the consumer and resource
species involved in each trophic interaction (i.e.~mean mass).
Furthermore, it incorporates information about the interaction itself,
such as the dimensionality (2 or 3 dimensions).

This marine food web compiles all the trophic data available for the
high Antarctic Weddell Sea collected since 1983, and is one of the most
highly-resolved marine food webs documented to date. It's noteworthy
that it is a summary network that ignores seasonal changes
\citep{Jacob2011}.

@@

To estimate the strength of each interaction in the food web we followed
the methodology proposed by \citet{Pawar2012}. The minimum data
requirements are: body mass of the consumer (predator) and resource
(prey), and the interaction dimensionality classified as 2 or 3
dimensions. GATEWAy v.1.0 does provide the mean mass for consumers and
resources (except for `detritus' and `sediment') and the dimensionality
for the majority of the interactions, though the latter is missing in
some cases (924 interactions). To complete such missing data, we used
information about movement type for consumer and resource included in
GATEWAy. Thus, we classified the interaction as 2D if both consumer and
resource move in 2D (e.g., both are sessile or walking) or if a consumer
moves in 3D and a resource in 2D (e.g., swimming consumer and
sessile/walking resource). The interaction was classified as 3D if both
consumer and resource move in 3D (e.g., both swimming) or if the
consumer moves in 2D and the resource in 3D (e.g., sessile/walking
consumer, swimming resource) \citep{Pawar2012}.

@@

We obtained estimations for the resource density and the search rate
from the scaling relationships with the resource and the consumer mass,
respectively \citep{Pawar2012}. The coefficients of such relationships,
determined by ordinary least squares regression, vary with the
interaction dimensionality. On one hand, resource density scales with
resource mass as power-law with exponents \vec{p = -0.79 \pm 0.09} in 2D
and \vec{p = -0.86 \pm 0.06} in 3D. Since mean mass for resources
`phytodetritus' and `sediment' were not available in GATEWAy, we
considered the body mass of the smallest phytoplankton species
(`Fragilariopsis cylindrus') as a proxy. This is justified by the fact
that `phytodetritus' and `sediment' are mainly composed by dead or
senescent phytoplankton reaching the seabed (\citet{Wolanski2011}). On
the other hand, search rate scales with consumer mass as power-law with
exponents \vec{p = 0.68 \pm 0.12} in 2D and \vec{p = 1.05 \pm 0.08} in
3D.

Finally, we fit the interaction strength distribution of the food web
considering six candidate models (Exponential, Gamma, log-Normal,
Normal, Power-law and Uniform) using maximum likelihood
\citep{McCallum2008}, and selected the model performance by computing
the Akaike Information Criterion AIC \citep{Burnham2002}.

@@

In order to individually characterize the species of the food web, we
considered weighted and unweighted properties (Figure 2). The former is
based on the estimation of the interaction strength described in the
previous section. The latter is related to properties commonly used in
qualitative or topological (presence/absence of interaction) food web
studies \citep{Martinez1991, Dunne2002, Borrelli2014}.

As the weighted property we took into account the total mean interaction
strength, meaning the average strength of all species' interactions for
each species. On the other hand, we considered the following unweighted
properties: a) degree or the total number of trophic interactions,
summing up in- and out-coming interactions (role as predator and prey,
respectively); b) trophic level or the position in the food web relative
to primary producers/detritus; and c) trophic similarity or the trophic
overlap based on shared and unique resources and consumers. The
following are arguments to have selected the unweighted properties. The
degree has often been equated with importance to the structure and
functioning of a community, i.e.~perturbations to high-degree species
may therefore have larger effects on the food web than perturbations to
low-degree species \citetext{\citealp{Dunne2002a}; \citealp[references
in][]{Cirtwill2018a}}. The trophic level offers information about how
important a species is to its biotic community, i.e.~top predators and
primary producers are expected to have particularly large effects on the
rest of their communities through top-down and bottom-up control,
respectively \citep[references in][]{Cirtwill2018a}. The trophic
similarity measures one of the most important aspects of species'
niches, the trophic niche, and functional aspects of biodiversity
\citep{Martinez1991, Williams2000}.

Furthermore, we took into account the species' habitat, which describes
the physical position of a species within the ecosystem. Species were
categorized as: 1) benthic, if the species lives on the seafloor; 2)
pelagic, if the species lives close to the surface; 3) benthopelagic, if
it moves between and connects the mentioned environments; 4) demersal,
if it lives and feeds on or near the bottom of the sea; and 5)
land-based, if the consumer is not aquatic but feeds predominantly in
the marine realm. These data was taken from \citet{Jacob2011}.

With the aim of studying the relationship between the interaction
strength of the species (weighted property) and its unweighted
properties we performed linear regression analyses between the log(mean
interaction strength) and each of the mentioned unweighted properties.
We also explored such weighted property with species habitat.

@@

Finally, we run extinction simulations and estimated its impact on the
stability of the food web. For this, we calculated a stability index
called Quasi-Sign Stability (QSS), which is the proportion of stable
networks using randomized Jacobians and keeping the predator-prey sign
structure fixed \citep{Allesina2008}. If this proportion is zero, then
one should take into account the real part of the maximum eigenvalue of
the Jacobian matrix, which is also a measure of stability
\citep{Grilli2016}.

With the aim of analysing the effect of each species on the food web's
stability we performed extinction simulations deleting one species at a
time, so the network size was reduced by one. After each species
extinction, we calculated the stability for the food web minus one
species (size = 489) and compared it with that of the whole network
(size = 490). We performed 1000 simulations for each species extinction
and obtained a mean QSS and maximum eigenvalue. At last we statistically
analysed such difference with an Anderson-Darling test considering a
p-value \textless{} 0.01 \citep{Scholz1987}. If the difference was
positive, then the stability of the food web was impacted positively
(higher stability) when that species become extinct, and viceversa.
Details for the stability calculations are described in Supplementary
Material.

Once we had the results for the impact on stability for each species
extinction, we plotted them considering weighted (interaction strength)
and unweighted properties, and species habitat. With this we aim to
characterize those species with a relatively high effect on the
stability of the food web.

All analyses were performed in R software, mainly using packages igraph
\citep{Csardi2005}, cheddar \citep{Hudson2013}, and multiweb
\citep{Saravia2019}. The source code and data are available at
https://github.com/EcoComplex/WeddellSea.

@@

\subsection{Interaction strength}

In this work we have estimated the interaction strength for the most
highly-resolved marine food web to date, which comprises 490 species and
16041 predator-prey interactions. The distribution of the interaction
strength best fit to a log-Normal model, meaning that there is a
prevalent skew towards weaker interactions (Figure 3, Table 1).

\subsection{Species properties}

Following the above results for the distribution of the interaction
strengths, we found that the species mean interaction strength was
different. We also found that such weighted property is related more or
less with unweighted properties (Figure 4A-D). In this regard,
interaction strength - trophic level showed a significant positive
linear regression: the higher the trophic level of the species, the
higher its mean interaction strength. We also found a significant but
less evident positive relationship with degree or the total number of
interactions. On the other hand, there was no significant relationship
with trophic similarity. Considering species habitat, the
``Benthopelagic'' and ``Pelagic'' categories presented the two species
with the highest mean interaction strength (Orcinus orca and
Mesonychoteuthis hamiltoni, respectively), though ``Demersal'' and
``Land-based'' habitats concentrated the majority of the species with
relatively higher interaction strength. Species inhabiting the benthic
realm showed the lowest mean interaction strength (Figure 4D).

\subsection{Extinction simulations and stability}

It's important to note that since the proportion of Jacobians that are
locally stable or QSS was zero for the Weddell Sea food web, we
considered the mean maximum eigenvalue as the stability index. When
performing the extinction simulations, our results showed that the
majority of the species had no significant impact on the stability of
the food web when extinct (Figure 5). This is shown in Figure 5, where
most of the points lie around the zero value of the `Stability
difference' y axis, meaning that the stability for the food web minus a
given species is similar (not significant) to the stability for the
whole network. However, there were few species that when become extinct
altered significantly the stability of the food web. Most of such
species generated positive significant differences in the stability of
the food web, which mean that they increased the network's stability
when deleted. There were two exceptions, Pagetopsis macropterus
(demersal fish) and Maxilliphimedia longipes (benthopelagic amphipod),
that when deleted decreased the stability. Overall, 15 out of 490
species (3.06\%) gave rise to significant changes in the food web's
stability (Table 2).

After exploring the stability difference against the species properties
(Figure 5), we found that those species that generated a significant
impact on the stability of the food web were characterized by: 1) high
mean interaction strength; 2) mid to high trophic levels (TL
\textgreater{} 3.2); 3) relatively high number of interactions (Degree
\textgreater{} 25); and 4) mid to low trophic similarity (TS \textless{}
0.16). Habitat wise, species with a significant impact on the stability
were present in all habitats, except for the benthic realm. Table 2
shows these results for such species.

@@

Using the highest-resolved marine polar food web up to date, we show
that 1) the distribution of the interaction strength at the food web
level is asymmetric, where weak interactions are prevalent; that 2)
there is a positive relationship between species mean interaction
strength and two unweighted properties, trophic level and degree; and
that 3) only few species are key in terms of food web stability,
presenting particular weighted and unweighted properties. It's also
noteworthy, that our study is the first of its type to consider the
dimensionality of the interaction (2 or 3 dimensions) to estimate
predator-prey relationships within a food web framework. After analysing
more than 2900 predator-prey interactions from several ecosystems,
\citet{Pawar2012} concluded that interaction dimensionality is a
critical factor driving consumer-resource dynamics, which will lead to
better predictions of food web and ecosystem functioning. This arises as
crucial to better understand how a complex system such as the Weddell
Sea might respond to environmental change, which is an ongoing issue in
that region of the Antarctic \citep[ and references herein]{Gutt2022}.

Our results regarding the distribution of the interaction strengths are
consistent with prior theoretical and empirical studies
\citep[e.g.][]{McCann1998, Neutel2002, Emmerson2004, Wootton2005}.
Within a system of a given complexity, commonly only a few strong
species interactions are present with most interactions being weak.
Historically, after \citet{May1972} installed the debate
complexity-stability, the results about the distribution of the
interaction strength has been interpreted as an explanation for the
persistence of complex food webs in nature
\citep{Bascompte2005, Allesina2015, Nilsson2016}. Here we showed that
this pattern is also prevalent in one of the most complex empirical
marine food webs that was studied so far, with 490 species and 16041
predator-prey interactions.

Our approach to study species properties follows \citet{Kortsch2021}
recommendation that suggests to employ a range of descriptors from both
unweighted and weighted food web in order to characterise the dynamic
and multifaceted nature of structural and functional ecosystem changes.
We went a step forward and analysed the relationship between weighted
and unweighted descriptors. In this regard, our results show a positive
relationship between interaction strength (weighted) and trophic level
(unweighted), and between interaction strength and degree (total number
of interactions). The former relationship might contradict those studies
that suggest that mid-trophic level species are involved in the major
pathways of energy flow in high-latitude marine ecosystems
\citep{Pinkerton2014, Murphy2016, McCormack2020, Riccialdelli2020}. This
could be explained by the lack of species biomass data in our
interaction strength estimations; the methodology we applied here
\citep{Pawar2012} allows to include empirical data for the density of
the resource, though not for all food web species. Unfortunately, the
lack of individual data for the entire Weddell Sea food web hampers any
alternative approach. On the other hand, the positive relationship
between interaction strength and degree reinforces the importance of
species with many interactions: species with high degree (hubs) have a
large impact on overall food web structure and functioning
\citep{Dunne2002a, Kortsch2015}. In the Weddell Sea, species with high
degree also tend to have high mean interaction strengths. This
information on the quantity and quality of interactions and its
relationship enables a robust assessment of the species' role in the
stability of the food web \citep{Cirtwill2018a}.

Our results show that only few species are key in terms of stability in
the Weddell Sea food web. This is in concordance with studies of complex
empirical food webs in marine ecosystems in the Arctic and other
locations in Antarctica \citep{Kortsch2015, Marina2018, Rodriguez2022}.
In the present work we identify that such key species present particular
weighted and unweighted properties: 1) high mean interaction strength;
2) mid to high trophic level; 3) relatively high number of interactions;
and 4) mid to low trophic similarity. The first attempt to gain insight
into the role of species in the Weddell Sea applying a complex approach
was that of \citet{Jacob2011}. These authors used information on body
size, predatory and feeding type, and environment (habitat) to study
species' roles. After performing sequential extinction simulations,
\citet{Jacob2011} found that larger bodied species can be lost without
causing a direct collapse of the network topology. A major caveat of
this finding, already recognized by the authors, is that population
dynamics were ignored and therefore no top-down extinctions, or other
indirect effects, could occur. In our study we did consider such effects
by including information on the species interaction strength, which is
of paramount importance when analysing the response of perturbations in
ecological communities \citep{McCann1998, Montoya2009, Novak2011}. Thus,
our study suggests that species with high mean interaction strength and
high trophic level need to be considered with particular attention when
trying to predict the effects of perturbations on the Weddell Sea
ecosystem. Furthermore, this arises as a greater issue since we have
identified that these species present a mid to low trophic similarity,
meaning that few other species of the food web can occupy the same
trophic role. In this sense, \citet{Murphy2016}'s review emphasized that
polar pelagic ecosystems are particularly sensitive to change due to a
low functional redundancy at key trophic levels. It's important to note
that we provide a broader analysis in this regard because we have
included species from all habitats (pelagic, benthic, land-based).

@@

Our study goes beyond the current understanding of how species influence
ecosystem structure and stability in the Weddell Sea in particular and
in most polar regions in general \citep{Murphy2016, McCormack2021}. In
the same analysis we have integrated information about weighted
(interaction strength) and unweighted species properties, enabling a
more complete assessment of the food web structure and function, likely
highlighting the ecological processes at play in the Weddell Sea
ecosystem.

@@

\begin{figure}
\includegraphics[width=12cm]{Fig3_IntDist} \caption{Frequency distribution of interaction strengths for the Weddell Sea food web. Total number of interactions = 16041.}\label{fig:unnamed-chunk-3}
\end{figure}

@@

\begin{figure}
\includegraphics[width=12cm]{Fig.5_QSSDif} \caption{Quasi-Sign Stability (QSS) difference between the whole Weddell Sea food web (n = 490) and the food web without one species (n = 489) for weighted (interaction strength) and unweighted species properties, and habitat. Point color indicates the impact on the QSS; if significant the extinction of that species altered the stability (QSS) of the food web.}\label{fig:unnamed-chunk-5}
\end{figure}

\begin{table}[t]
\caption{Model comparison for the distribution of interaction strengths of the Weddell Sea food web. Order by best fit. References: df = degrees of freedom, AIC = Akaike Information Criterion, deltaAIC = difference with best fit. Log-Normal is the best model.}
\begin{tabular}{l c c c}
\tophline

\textbf{Model} & \textbf{df} & \textbf{AIC} & \textbf{deltaAIC} \\
\middlehline
log-Normal & 2 & -359738 & 0 \\
\middlehline
Gamma & 2 & -359714.1 & 23.89 \\
\middlehline
Power-law & 2 & -350667 & 9070.97 \\
\middlehline
Exponential & 1 & -327606.5 & 32131.54 \\
\middlehline
Normal & 2 & -291407.8 & 68330.21 \\
\middlehline
Uniform & 2 & -248179 & 111559.02 \\

>>>>

\begin{abstract}
Network approaches can shed light on the structure and stability of
complex marine communities. In recent years, such approaches have been
successfully applied to study polar ecosystems, improving our knowledge
on how they might respond to ongoing environmental changes. The Weddell
Sea is one of the most studied marine ecosystems outside the Antarctic
Peninsula in the Southern Ocean. Yet, few studies consider the known
complexity of the Weddell Sea food web which in its current form
comprises 490 species and 16041 predator-prey interactions. Here we
analysed the Weddell Sea food web, focusing on trophic interactions that
underpin ecosystem structure and stability. We estimated the strength
for each interaction in the food web, characterised species position in
the food web using unweighted and weighted properties, and analysed
species' roles with respect to the stability of the food web. On one
hand, we found that the distribution of the interaction strength at the
food web level is asymmetric, where weak interactions are prevalent.
Including such information as a (weighted) property for species we
detected a positive relationship between species mean interaction
strength and two unweighted properties, trophic level and the total
number of interactions. We also found that only a few species are key in
terms of food web stability, presenting high mean interaction strength,
mid to high trophic level, relatively high number of interactions, and
mid to low trophic similarity. In the same analysis we have integrated
food web and species information, enabling a more complete assessment of
the ecosystem structure and function, likely highlighting the ecological
processes at play in the Weddell Sea. We consider that our results
provide new insights important for the development of effective policies
and management strategies, particularly given the ongoing initiative to
implement a Marine Protected Area (MPA) in the Weddell Sea.
\end{abstract}

@@

The objective of this work was to improve the knowledge on how the
Weddell Sea and the species therein may respond to perturbations from
ongoing environmental changes. To achieve this we: 1) estimated the
strength for each interaction in the Weddell Sea food web, 2)
characterised species considering weighted and unweighted properties,
and 3) analysed the species' role in the stability of the food web.

@@

The high Antarctic Weddell Sea shelf is situated between 74 and 78ºS
stretching approximately 450 km from East to West (Figure 1). Water
depth varies between 200 and 500 meters, and shallower areas are covered
by continental ice, which forms the coastline along the eastern and
southern part of the Weddell Sea. The shelf area contains a complex
benthic three-dimensional habitat with large benthic biomasses,
intermediate to high diversity in comparison to benthic boreal
communities and a spatially patchy distribution of organisms
\citep{Dayton1990, Teixido2002}.

@@

The Weddell Sea food web was retrieved from the GlobAL daTabasE of
traits and food Web Architecture (GATEWAy, version 1.0) of the German
Centre for Integrative Biodiversity Research (iDiv) Halle-Jena-Leipzig
\citep{Brose2018}. In addition to predator-prey interactions, the
database contains information on other biological data such as the mean
body mass and movement type for each species in the food web.
Furthermore, it incorporates information about the interaction itself,
such as the dimension of the predator search space (2 or 3 dimensions).
In its current form the Weddell Sea food web comprises 490 species and
16041 predator-prey interactions and constitutes one of the most
resolved food webs constructed to date \citep{Jacob2011}.

@@

To estimate the strength of each pairwise interaction in the food web we
followed the methodology proposed by \citet{Pawar2012}. The minimum data
requirements are body mass of the consumer (predator) and resource
(prey), and the interaction dimensionality (ID) classified as 2 or 3
dimensions. The ID is defined as the dimension of the search space of
the predator, that is equivalent to the movement space of the prey.
Thus, the ID is classified as 2D when both predator and prey move in 2D
(e.g., both are benthic) or if a predator moves in 3D and a prey in 2D
(e.g., pelagic predator on benthic prey). The ID is classified as 3D
when both predator and prey move in 3D (e.g., both pelagic) or if the
predator moves in 2D and the prey in 3D (e.g., benthic predator, pelagic
prey) \citep{Pawar2012}. GATEWAy v.1.0 provides information on the mean
body mass for consumers and resources, except for `detritus' and
`sediment', and the dimensionality for the majority of the interactions,
though the latter is missing in some cases (924 interactions). To
complete the missing data on species `dimensionality', we used
information about the movement type of predators and prey included in
GATEWAy.

@@

We obtained estimates for resource density and the search rate from the
scaling relationships with the resource and the consumer mass,
respectively \citep{Pawar2012}. The coefficients of such relationships,
determined by ordinary least squares regression, vary with the
interaction dimensionality. On one hand, resource density scales with
resource mass as power-law with exponents in 2D and in 3D. Since mean
mass for resources `phytodetritus' and `sediment' were not available in
GATEWAy, we considered the body mass of the smallest phytoplankton
species (`Fragilariopsis cylindrus') as a proxy. This is justified by
the fact that `phytodetritus' and `sediment' are mainly composed of dead
or senescent phytoplankton reaching the seabed \citep{Wolanski2011}. On
the other hand, search rate scales with consumer mass as power-law with
exponents in 2D and in 3D.

We fitted six candidate models (Exponential, Gamma, log-Normal, Normal,
Power-law and Uniform) the interaction strength distribution using
maximum likelihood \citep{McCallum2008}, and selected the best fitting
model by computing the Akaike Information Criterion AIC
\citep{Burnham2002}.

@@

To characterise the role of each species in the food web, we considered
unweighted and weighted food web properties (Figure 2). Unweighted
properties are related to properties commonly used in qualitative food
web studies and only describe the presence or absence of interactions
without any information on strength between a pairwise species link
\citep{Martinez1991, Dunne2002, Borrelli2014}. In contrast, weighted
properties capture the importance of interaction strength.

To assess species roles as a function of the weighted food web, we
focused on mean interaction strength defined as the average strength of
all interactions for a given species. Further we calculated three
unweighted species properties: a) species degree, i.e., the sum of in-
and out-going interactions ; b) trophic level ; and c) trophic
similarity, i.e., the trophic overlap based on shared and unique
resources and consumers. These metrics were chosen to assess a species
role based on the unweighted food web. The species degree has often been
equated with species importance to the structure and functioning within
a food web, i.e.~perturbations to high-degree species may therefore have
more significant effects on the food web robustness to perturbations
than low-degree species
\citetext{\citealp{Dunne2002a}; \citealp[references
in][]{Cirtwill2018a}}. The trophic level offers information about how
important a species is to its biotic community, i.e., top predators and
primary producers are expected to have particularly large effects on the
rest of their communities through top-down and bottom-up control,
respectively \citep[references in][]{Cirtwill2018a}. Trophic similarity
is an index of trophic overlap considering the set of prey and predators
for a pair of species; it measures one of the most important aspects of
species' niches, the trophic niche, and functional aspects of
biodiversity \citep{Martinez1991, Williams2000}.

Furthermore, we took a species' habitat into account, which describes
the physical position of a species within the ecosystem. Species were
categorised as: 1) benthic, if a species lives on the seafloor; 2)
pelagic, if a species lives close to the surface; 3) benthopelagic, if
it moves between and connects the mentioned environments; 4) demersal,
if it lives and feeds on or near the bottom of the sea; and 5)
land-based, if the consumer is not strictly aquatic but feeds
predominantly on marine species. Species habitat affiliation was
retrieved from \citet{Jacob2011}.

With the aim of studying the relationship between the interaction
strength of the species (weighted property) and its unweighted
properties we performed linear regression analyses between the log mean
interaction strength and each of the mentioned unweighted properties.
Thus we considered the interaction strength as the dependent variable
and the given unweighted property as the independent variable, and
obtained the coefficients (slope and intercept) for the linear model.
Models were fitted using the least squares approach. We also explored
the mean interaction strength distribution with the species habitat.

@@

To analyse the impact of species on food web stability, we performed
extinction simulations deleting one species at a time, that is for every
extinction network size was reduced by one species only. After each
extinction, we calculated the stability of the network minus the
targeted species (489 nodes) and compared it with that of the whole
network (490 nodes in total). To calculate stability, we used the mean
of the real part of the maximum eigenvalue of the Jacobian matrix using
randomized Jacobians and keeping the predator-prey sign structure fixed
\citep{Allesina2008, Grilli2016}. This index indicates a more stable
food web when it is negative. We performed 1000 simulations for each
species extinction and obtained a mean maximum eigenvalue for each case.
At last we statistically analysed such a difference with an
Anderson-Darling test considering a p-value \textless{} 0.01
\citep{Scholz1987}. If this difference is positive, then the stability
of the food web is higher if a species was removed, and vice versa. A
detailed description on the stability calculations can be found in
Supplementary Material.

Once we had the results for the impact on stability for each species
extinction, we plotted them considering weighted (interaction strength)
and unweighted properties, and species habitat. With this we aim to
characterise those species with a relatively high effect on the
stability of the food web.

All analyses were performed in R software, using the R packages igraph
\citep{Csardi2005}, cheddar \citep{Hudson2013}, and multiweb
\citep{Saravia2019}. The source code and data are available at
https://github.com/EcoComplex/WeddellSea.

@@

\subsection{Interaction strength distribution}

The statistical distribution that best fitted the empirical interaction
strength distribution was a `gamma' due to the high proportion of weak
interactions and the existence of a few strong interactions (Figure 3,
Table S3).

\subsection{Species' role related to their mean interaction strength}

We found that the species' mean interaction strength (weighted property)
shows different relationships with the unweighted properties analysed
(Figure 4A-D). In this regard, there is a positive relationship between
interaction strength and trophic level, i.e., the higher the trophic
level of the species, the higher its mean interaction strength. We also
found a significant but less evident positive relationship with species
degree. Contrary, there was no significant relationship between mean
interaction strength and trophic similarity. Considering species habitat
affiliation, the ``Benthopelagic'' and ``Pelagic'' categories contained
the two species with the highest mean interaction strength, the killer
whale Orcinus orca and the colossal squid Mesonychoteuthis hamiltoni,
respectively. However, the majority of the species with relatively
higher interaction strength belonged to the ``Demersal'' and
``Land-based'' habitats groups. Species inhabiting the benthic realm
showed the lowest mean interaction strength (Figure 4D).

\subsection{Species impact on food web stability}

Our extinction analyses showed that the majority of species had no
significant impact on food web stability after being removed (Figure 5).
Most of the species (black points in figure 5) did not change the
stability of the network considerably after being removed, except for a
few species. Only 15 out of 490 species (3.06\%) gave rise to
significant changes in the food web's stability after their removal
(Table 2). Most of these species had a positive impact on food web
stability, i.e., network stability increased after their removal. Only
two species significantly decreased network stability after being
removed, the demersal fish Pagetopsis macropterus and the benthopelagic
amphipod Maxilliphimedia longipes.

After exploring the stability difference against the species properties
(Figure 5), we found that those species that generated a significant
impact on the stability of the food web were characterised by: 1) high
mean interaction strength; 2) mid to high trophic levels (TL
\textgreater{} 3.2); 3) relatively high number of interactions (Degree
\textgreater{} 25); and 4) mid to low trophic similarity (TS \textless{}
0.16). Habitat wise, species with a significant impact on the stability
were present in all habitats, except for the benthic realm. Table 2
shows these results for such species.

@@

\subsection{Many weak and a few strong interactions}

Our analyses show that the distribution of species interaction strength
at the network level is asymmetric, i.e., the Weddell Sea food web
contains many weak interactions and only a few strong ones. This finding
is consistent with many previous theoretical and empirical studies
\citep[e.g.][]{McCann1998, Neutel2002, Emmerson2004, Wootton2005, Kortsch2021}.
The asymmetric distribution of interaction strength in food webs has
been interpreted as an explanation for the persistence of complex
communities in nature \citep{Bascompte2005, Allesina2015, Nilsson2016}.
Here we show that this pattern is also prevalent in one of the most
complex empirical (marine) food webs to date, comprising 490 species and
16041 predator-prey interactions. This finding reinforces the call for
the inclusion of interaction strength in food web studies to better
understand the ecosystem functioning, and species and whole network
responses to environmental perturbations.

\subsection{Species's role related to their mean interaction strength}

We employed a range of descriptors using both unweighted and weighted
food web properties to characterise the dynamic and multifaceted nature
of the Weddell Sea food web. Our results show a positive relationship
between interaction strength and trophic level, and between interaction
strength and species degree. Mean interaction strength increases with
trophic level and species degree. The former relationship might
contradict those studies that suggest that mid-trophic level species are
involved in the major pathways of energy flow in high-latitude marine
ecosystems
\citep{Pinkerton2014, Murphy2016, McCormack2020, Riccialdelli2020}. This
could be explained by the lack of species biomass data in our
interaction strength estimations; the methodology we applied here
\citep{Pawar2012} allows empirical data for the density of the resource
to be included, though this data is not for the majority of food web
species at the study site. On the other hand, the positive relationship
between interaction strength and degree reinforces the importance of
species with many interactions: species with high degree (hubs) have a
large impact on overall food web structure and functioning
\citep{Dunne2002a, Kortsch2015}. In the Weddell Sea, species with high
degree also tend to have high mean interaction strengths. This
information on the quantity and quality of interactions and its
relationship enables a robust assessment of the species' role in the
stability of the food web \citep{Cirtwill2018a}.

\subsection{Species impact on food web stability}

Only a few species play a key role with respect to the Weddell Sea food
web stability, according to the mean maximum eingenvalue stability index
employed in this study. This is in concordance with other studies on
complex empirical food webs in marine ecosystems in the Arctic and other
locations in Antarctica \citep{Kortsch2015, Marina2018, Rodriguez2022}.
These key species are characterised by a particular set of food web
properties: high to mean interaction strength; mid to high trophic
level; relatively high number of interactions; and mid to low trophic
similarity. In a previous study on sequential extinction simulations for
the Weddell Sea food web, it was found that larger bodied-sized species
could be lost without causing a collapse of the network. A major caveat
of this finding, also recognised by the authors, was that population
dynamics were ignored and hence no top-down extinctions, or other
indirect effects, could occur. In our study we considered such top-down
effects by including information on the species interaction strength,
which is of paramount importance when analysing the response of
perturbations in ecological communities
\citep{McCann1998, Montoya2009, Novak2011}. Thus, our study suggests
that species with high mean interaction strength and high trophic level
need to be considered with particular attention when trying to predict
the effects of perturbations on the Weddell Sea ecosystem. This
conclusion is further reinforced by the finding that these species have
mid to low trophic similarity, which means that few other species of the
food web can occupy the same trophic role. In a review, it was
emphasised that polar pelagic communities are particularly sensitive to
changes due to a low functional redundancy at key trophic levels
\citep{Murphy2016}. Here we provide broader analyses of species impact
on food web robustness by including species from all habitats (benthic,
pelagic and land-based). This suggests that the sensitivity of marine
polar ecosystems to environmental perturbations is a concern also beyond
the pelagic realm.

@@

Our study goes beyond the current understanding of how species influence
ecosystem structure and stability in the Weddell Sea in particular and
in most polar regions in general \citep{Murphy2016, McCormack2021}. In
the same analysis we have integrated information about weighted
(interaction strength) and unweighted species properties, enabling a
more complete assessment of the species' role in the food web structure
and function, likely highlighting the ecological processes at play in
the Weddell Sea ecosystem.

@@

\begin{figure}
\includegraphics[width=12cm]{Fig3_IntDist} \caption{Frequency distribution of interaction strengths for the Weddell Sea food web. Total number of interactions = 16041. The distribution was best fitted to a ‘gamma’ model.}\label{fig:unnamed-chunk-3}
\end{figure}

@@

\begin{figure}
\includegraphics[width=12cm]{Fig.5_QSSDif} \caption{Stability  difference (mean maximum eingenvalue) between the whole Weddell Sea food web (n = 490) and the food web minus one species (n = 489) for weighted (interaction strength) and unweighted species properties, and habitat. Point color indicates the impact on the stability; if significant the extinction of that species altered the stability of the food web.}\label{fig:unnamed-chunk-5}
\end{figure}
